% Chapter 1

\chapter{Machine learning for an IDS} % Main chapter title

\label{Chapter3} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Using ML for an IDS}
An intrusion detection system has to detect whether some data it receives is either malicious or regular web traffic. This can be seen as a classification problem which means an machine learning algorithm for classification could be used. It needs to be determined whether data is either normal network traffic or malicious behaviour. \\
\\
Some parameters have to be chosen that will be feed into the machine learning algorithm.

\section{Disadvantages of using ML for an IDS}
\subsection{Problems}
As said before, machine learning for an intrusion detection system is a classification problem. More precisely, it can be said that intrusion detection systems have to detect abnormal behaviour in a network with mostly normal behaviour. There are several problems that can be encountered when using machine learning techniques.\\
\\
The first problem is the ability to detect new attacks. A machine learning algorithm compares incoming data with a model that it has created internally. An new type of malicious behaviour might appear to be closer to normal network traffic as compared to the model of known attacks. \\
\\
Another problem is the diversity of network traffic. The notion of "normal network traffic" is difficult to actually define. The bandwidth, duration of connections, origin of IP addresses, applications used can vary enormously through time. This makes it quite difficult for machine learning algorithms to distinguish between "normal network traffic" and malicious behaviour.\cite{ClosedWorld} 

\subsection{Solutions}
There are several solutions that can be used in order to make machine learning algorithms more effective for intrusion detection systems. One option is to chance the way the classification problem is defined. Instead of defining the classes, "normal" and "malicious", there might be different classes for different types of malicious behaviour. In the same way, different classes can be defined for different types normal traffic. 

\section{Advantages of using ML for an IDS}

